---
title: It's Getting *Hot* in Here
subtitle: Analysis of Global CO2 Emissions Over Time
author: Impressive Squirtle <br> Cassidy Cheesman, Joanne Wang, Ming DeMers, Elisabeth Pan, and Ivan Lin
date: 5/5/23
format: 
  revealjs:
    smaller: true
editor: visual
execute:
  echo: false
---

## Topic and Motivation

-   Climate change is the biggest challenge facing our generation, with increased global CO2 emissions being one of the leading contributors to this issue.

-   In order to combat the effects of climate change, we must first understand the data behind it.

-   Research questions:

    -   What is the relationship between GDP and CO2 emissions for developing versus developed countries in 2018?

    -   Is the true proportion of countries with a CO2 emission greater than the US in the twentieth century different from those in the twenty-first century?

## The Data

-   Created to show how emissions are changing in each country

-   The Our World in Data team created the original dataset on behalf of the Our World in Data project

-   The instances that comprise the dataset represent countries (per year from 1992 to 2018) with 2484 instances total.

-   The observations (rows) are countries per year and the attributes (columns) are various emission data for each country per year

-   The current dataset we will use is a subset of the original dataset, with all the tonnes units converted into kilotons and the variable names altered

## Highlights from the EDA

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
library(tidymodels)
library(skimr)
library(rcis)
```

```{r}
#| label: read-csv-glimpse

pollution <- read_csv("data/global_emissions.csv")
```

::: columns
::: {.column width="33%"}
```{r}
#| label: gdp-developed
#visualizing change in GDP over time for a developed country
pollution_usa <- pollution |>
  filter(Country.Name == "United States")

ggplot(data = pollution_usa, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in USA", 
    x = "Year",
    y = "GDP (international$)"
  )
```
:::

::: {.column width="33%"}
```{r}
#| label: gdp-developing
#visualizing change in GDP over time for a developing country
pollution_pakistan <- pollution |>
  filter(Country.Name == "Pakistan")

ggplot(data = pollution_pakistan, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in Pakistan", 
    x = "Year",
    y = "GDP (international$)"
  ) 
```
:::

::: {.column width="33%"}
```{r}
#| label: CO2-developed
#visualizing change in CO2 emissions over time for a developed country
ggplot(data = pollution_usa, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Total)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in USA", 
    x = "Year",
    y = "Total CO2 Emissions (kilotons)"
  ) 
```
:::
:::

::: columns
::: {.column width="33%"}
```{r}
#| label: CO2-developing
#visualizing change in CO2 emissions over time for a developing country
ggplot(data = pollution_pakistan, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Total)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in Pakistan", 
    x = "Year",
    y = "Total CO2 Emissions (kilotons)"
  ) 
```
:::

::: {.column width="33%"}
```{r}
#| label: gdp-CO2-1992
#visualizing relationship between GDP and total CO2 emissions in 1992
pollution_92 <- pollution |>
  filter(Year == 1992)

ggplot(data = pollution_92,
       mapping = aes(x = Country.GDP, y = Emissions.Production.CO2.Total)
       ) + 
  geom_point(alpha = 0.5) + 
  xlim(0, 3e+12) + 
  ylim(0, 2000) + 
  theme_minimal() + 
  labs(
    title = "Total CO2 Emissions vs. Country GDP in 1992", 
    x = "GDP (international$)",
    y = "Total CO2 Emissions (kilotons)"
  ) 
```
:::

::: {.column width="33%"}
```{r}
#| label: gdp-CO2-2018
#visualizing relationship between GDP and total CO2 emissions in 2018
pollution_18 <- pollution |>
  filter(Year == 2018)

ggplot(data = pollution_18,
       mapping = aes(x = Country.GDP, y = Emissions.Production.CO2.Total)
       ) + 
  geom_point(alpha = 0.5) + 
  xlim(0, 4e+12) + 
  ylim(0, 2000) + 
  theme_minimal() + 
  labs(
    title = "Total CO2 Emissions vs. Country GDP in 2018", 
    x = "GDP (international$)",
    y = "Total CO2 Emissions (kilotons)"
  )
```
:::
:::

## Inference, Modeling, and Analysis {.smaller}

Research question: What is the association between GDP and CO2 emissions for developing versus developed countries in 2018?

```{r}
developed_countries <- c(
  "Austria", "Belgium", "Denmark", "Finland", "France", "Germany", "Greece", "Ireland", "Italy", "Luxembourg", "Netherlands", "Portugal", "Spain", "Sweden", "United Kingdom", "Bulgaria", "Croatia", "Cyprus", "Czech Republic", "Estonia", "Hungary", "Latvia", "Lithuania", "Malta", "Poland", "Romania", "Slovakia", "Slovenia", "Iceland", "Norway", "Switzerland", "Australia", "Canada", "Japan", "New Zealand", "United States"
  )

pollution_18 <- pollution_18 |>
  mutate(developed = if_else(Country.Name %in% developed_countries, "Yes", "No"),
         developed = factor(developed))

log_fit <- logistic_reg() |>
  fit(developed ~ Emissions.Production.CO2.Total, data = pollution_18)

tidy(log_fit)

```

$$ log(\frac{p}{1-p})=-5.14 \times 10^{-1} +(- 5.69 \times 10^{-5} \times total~CO2~emissions) $$ The intercept can be interpreted as the log odds of a country being "developed" if total CO2 emissions is 0, which is $-5.14 \times 10^{-1}$. However, because we do not use or think in terms of log odds, we can further interpret this intercept by evaluating p. The odds, $\frac {p}{1-p}$, of a country being "developed" if total CO2 emissions is 0 is $e ^ {-5.14 \times 10 ^ {-1}}$ which is 0.598, where p represents the probability that the country is developed. Thus, given that total CO2 emissions is 0, the probability a country is developed is $e ^ {-5.14 \times 10 ^ {-1}} - e ^ {-5.14 \times 10 ^{-1}} p$ or 0.37.

##  {background-video="recorded-slide.mp4"}

## Conclusions and Future Work

-   We find that there is indeed an increase in CO2 emissions between the two centuries

-   For developing countries, a near-exponential increase

-   For developed countries, a linear increase.

-   This requires transitioning to renewable energy sources, improving energy efficiency, reducing deforestation, and promoting reforestation and eco-repair.

-   Not all 195 countries, nor only 1992 - 2018.

-   In the future, we hope to explore a larger time frame, and examine more features than just GDP, such as population, industrialization, number of cities, etc.

-   Developing or developed, every nation has a duty to help protect and preserve the planet in which we all reside.
